---
title: "Problem With rJava On Mac OS X El Capitan"
author: "Peter von Rohr"
date: "5 January 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r PostGlobals, echo=FALSE,results='hide'}
sPostRootDir <- "/Users/peter/Data/Projects/GitHub/charlotte-ngs.github.io"
sImgRootDir <- file.path(sPostRootDir, "img")
sPostName <- "2016-01-05-MacOsXrJavaProblem"
sImgRootUrl <- paste("http://charlotte-ngs.github.io/img", sPostName, sep = "/")
sGlobalPostStatus <- "publish"
#sGlobalPostStatus <- "draft"

```


```{r PostDeployment, eval=FALSE, echo=FALSE,results='hide'}
# this should only be run at the end when everything is complete
deployPost <- function(psPostName, psPostFrom = ".",  psPostTo = file.path(sPostRootDir,"_posts")) {
  ### # do nothing, if post file does not exist
  sPostFile <- paste0(psPostName, ".md")
  sPostFn <- file.path(psPostFrom, sPostFile)
  if (file.exists(sPostFn))
    file.rename(from = sPostFn, to = file.path(psPostTo, sPostFile))
}
deployPost(psPostName = sPostName)
```

```{r ImageDeployment, echo=FALSE,results='hide'}
deployImg <- function(sImgPattern = "png", sImgFrom = ".", sImgTo = file.path(sImgRootDir,sPostName)) {
  ### # check that sImgTo is not null
  stopifnot(!is.null(sImgTo))
  ### # in case image destination directory does not exist, create it
  if (!dir.exists(sImgTo)) dir.create(sImgTo, recursive = TRUE)
  ### list of image files
  vImgFiles <- list.files(path = sImgFrom, pattern = sImgPattern)
  ### loop over files and deploy them
  for (f in vImgFiles) {
    file.copy(from = f, to = sImgTo, overwrite = TRUE)
  }
}
deployImg()
```


```{r ImagePath, echo=FALSE, results='hide'}
imagePath <- function(psPostStatus, plImageOptions){
  cat("![", 
      plImageOptions[[psPostStatus]]$AltImageName, 
      "](",
      plImageOptions[[psPostStatus]]$ImageUrl, 
      ")\n",
    sep = "")
}
```

Since I upgraded my Mac to OS X El Capitan (version 10.11.2) I am having problems with the R-package [`rJava`](https://www.rforge.net/rJava). From what I understood, eventually, it is not a problem of `rJava` or R or RStudio in the first place, but newer versions of Mac OS X do not play well with the latest versions of Oracle's Java (jdk). Furthermore, some new security features in Mac OS X prevent that environment path settings are passed along to applications like RStudio. 

In what follows, I listed the steps that solved for me the problem that the package `rJava` could not be loaded from within RStudio under Mac OS X El Capitan (version 10.11.2). 

### Problem statement
Unfortunately, the package `rJava` does not just fail to load, but the R GUI for Mac terminates instantly and in RStudio the R-session gets closed. Both effects are pretty annoying! See the screen shot below on how that might look like.

```{r ImageRsessionAbort, echo=FALSE, results='asis'}
sAltImgName <- "R session abort"
sImgFn <- "RSessionAbort.png"
imagePath(psPostStatus = sGlobalPostStatus,
          plImageOptions = list(draft = list(AltImageName = sAltImgName, 
                                             ImageUrl = sImgFn),
                                publish = list(AltImageName = sAltImgName, 
                                             ImageUrl = paste(sImgRootUrl, sImgFn, sep = "/"))))
```

### Some research
After searching the web, the details of the problem are still not clear to me. But it appears that OS X has problems recognizing newer versions Oracle's Java. OS X seams to be looking for a legacy version 6 of Apple's own build of Java. As shown in the next section one solution seams to be the installation of Apple's legacy version 6 of Java and a newer version (e.g. 1.8.0_66) of Oracle's Java at the same time. 

### The solution
The post available at https://github.com/snowflakedb/dplyr-snowflakedb/wiki/Configuring-R-rJava-RJDBC-on-Mac-OS-X solved my problem with `rJava` on Mac OS X 10.11.2. 

### Java 6
I already had XCode command line tools installed, so I started at section `Java 6`. I downloaded the legacy Java 6 runtime from https://support.apple.com/kb/DL1572?locale=de_DE and installed the downloaded `.dmg` file.

### Java 8
Since I had already installed the latest version of Oracle's jdk from http://www.oracle.com/technetwork/java/javase/downloads/jdk8-downloads-2133151.html, I skipped the step.

The output of `java_home` on my machine looks as follows

```
YARLI:Downloads peter$ /usr/libexec/java_home -V
Matching Java Virtual Machines (5):
    1.8.0_66, x86_64:	"Java SE 8"	/Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home
    1.7.0_67, x86_64:	"Java SE 7"	/Library/Java/JavaVirtualMachines/jdk1.7.0_67.jdk/Contents/Home
    1.7.0_25, x86_64:	"Java SE 7"	/Library/Java/JavaVirtualMachines/jdk1.7.0_25.jdk/Contents/Home
    1.6.0_65-b14-468, x86_64:	"Java SE 6"	/Library/Java/JavaVirtualMachines/1.6.0.jdk/Contents/Home
    1.6.0_65-b14-468, i386:	"Java SE 6"	/Library/Java/JavaVirtualMachines/1.6.0.jdk/Contents/Home

/Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home
```

Please note the 1.7 version are really not needed here.

### R for OS X
The latest version of R was already running, so I skipped this step. The check with `javareconf` gave the following result.

```
YARLI:Downloads peter$ R CMD javareconf
Java interpreter : /usr/bin/java
Java version     : 1.8.0_66
Java home path   : /Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home/jre
Java compiler    : /usr/bin/javac
Java headers gen.: /usr/bin/javah
Java archive tool: /usr/bin/jar
Non-system Java on OS X

trying to compile and link a JNI program 
detected JNI cpp flags    : -I$(JAVA_HOME)/../include -I$(JAVA_HOME)/../include/darwin
detected JNI linker flags : -L$(JAVA_HOME)/lib/server -ljvm
clang -I/Library/Frameworks/R.framework/Resources/include -DNDEBUG -I/Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home/jre/../include -I/Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home/jre/../include/darwin -I/usr/local/include -I/usr/local/include/freetype2 -I/opt/X11/include    -fPIC  -Wall -mtune=core2 -g -O2  -c conftest.c -o conftest.o
clang -dynamiclib -Wl,-headerpad_max_install_names -undefined dynamic_lookup -single_module -multiply_defined suppress -L/Library/Frameworks/R.framework/Resources/lib -L/usr/local/lib -o conftest.so conftest.o -L/Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home/jre/lib/server -ljvm -F/Library/Frameworks/R.framework/.. -framework R -Wl,-framework -Wl,CoreFoundation


JAVA_HOME        : /Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home/jre
Java library path: $(JAVA_HOME)/lib/server
JNI cpp flags    : -I$(JAVA_HOME)/../include -I$(JAVA_HOME)/../include/darwin
JNI linker flags : -L$(JAVA_HOME)/lib/server -ljvm
Updating Java configuration in /Library/Frameworks/R.framework/Resources
Done.
```

### rJava
As the version of `rJava` from CRAN did not work, I took the latest version from `rforge` and installed that one.

```
curl -#ROL https://www.rforge.net/rJava/snapshot/rJava_0.9-8.tar.gz
R CMD INSTALL rJava_0.9-8.tar.gz
```

A quick check whether `rJava` can be loaded, is done as follows

```
R --quiet -e 'library("rJava"); .jinit(); .jcall("java/lang/System", "S", "getProperty", "java.runtime.version")'
> library("rJava"); .jinit(); .jcall("java/lang/System", "S", "getProperty", "java.runtime.version")
[1] "1.8.0_66-b17"
```

### RStudio
Running the above check on the RStudio console still gives the following error

```
library("rJava"); .jinit(); .jcall("java/lang/System", "S", "getProperty", "java.runtime.version")
Error : .onLoad failed in loadNamespace() for 'rJava', details:
  call: dyn.load(file, DLLpath = DLLpath, ...)
  error: unable to load shared object '/Library/Frameworks/R.framework/Versions/3.2/Resources/library/rJava/libs/rJava.so':
  dlopen(/Library/Frameworks/R.framework/Versions/3.2/Resources/library/rJava/libs/rJava.so, 6): Library not loaded: @rpath/libjvm.dylib
  Referenced from: /Library/Frameworks/R.framework/Versions/3.2/Resources/library/rJava/libs/rJava.so
  Reason: image not found
Error: package or namespace load failed for ‘rJava’
```

For some reason which might be connected to a new security feature of Mac OS X, the file `libjvm.dylib` cannot be found. Loading this library explicitly, as was described in http://stackoverflow.com/questions/30738974/rjava-load-error-in-rstudio-r-after-upgrading-to-osx-yosemite, finally solves the problem. 

```{r CheckRJava}
dyn.load('/Library/Java/JavaVirtualMachines/jdk1.8.0_66.jdk/Contents/Home/jre/lib/server/libjvm.dylib')
require(rJava)
.jinit()
.jcall("java/lang/System", "S", "getProperty", "java.runtime.version")
```

```{r}
sessionInfo()
```



