---
title: "Insert An Image Into A Plot"
author: "Peter von Rohr"
date: "21. April 2015"
output:
  pdf_document:
    keep_tex: yes
---


## Basic Version
The purpose of this document is to show some possibilities of how to include an image into a ggplot. The material shown here is based on http://stackoverflow.com/questions/9917049/inserting-an-image-to-ggplot2. 

The basic example is to include the R-logo in the background of a ggplot. 


```{r}
### # installing required packages
for (p in c("png","grid","ggplot2")) {
  if (!is.element(p, installed.packages())) {
    install.packages(p)
  }
}
### # loading packages
library(png)
library(grid)
library(ggplot2)
### # read png image
img <- readPNG(system.file("img", "Rlogo.png", package="png"))
g <- rasterGrob(img, interpolate=TRUE)
### # draw the qplot
qplot(1:10, 1:10, geom="blank") +
  annotation_custom(g, xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
  geom_point()

```


## Smaller Image
The image can also be insterted as a smaller version at a given position specified by the position arguments to `annotation_custom`

```{r}
### # draw the qplot
qplot(1:10, 1:10, geom="blank") +
  annotation_custom(g, xmin=2, xmax=5, ymin=7, ymax=10) +
  geom_point()

```

## More elaborate stuff
On the site http://stackoverflow.com/questions/12463691/inserting-an-image-to-ggplot-outside-the-chart-area there is a more elaborate example