---
title: "Create A Blog Post on Github Pages"
author: "Peter von Rohr"
date: "24. August 2015"
output: html_document
---


## Abstract
This document shows how a blog post on github pages can be created. 


## Introduction
[Github pages](https://pages.github.com/) are a platform for blog posts which uses github as 
repository for source files. Source files are written in markdown and [jekyll](http://jekyllrb.com/) is used to translate markdown sources into static html pages. 


## Aim
We want to design a function that sets up a blog post. The function takes as input, the root directory for the blog posts, the name of the source file, the title and the author of the post.


## A First Trial
The function is called `createGhPagesBlogPost()` and looks as follows. 

```{r CreateGhPagesBlogPost}
createGhPagesBlogPost <- function(psBlogTitle, 
                                  psBlogAuthor, 
                                  psBlogName, 
                                  psRootDir = getwd(),
                                  psImgDir = NULL){
  sBlogFileStem <- paste(Sys.Date(), "-", psBlogName, sep = "", collapse = "")
  sBlogFileName <- paste(sBlogFileStem, ".md", sep = "", collapse = "")
  sBlogFileDir <- file.path(psRootDir, "_posts")
  sBlogFilePath <- file.path(sBlogFileDir, sBlogFileName)
  ### # if the blog post file already exists, we are doing nothing
  if (!file.exists(sBlogFilePath)) {
    cat("---\ntitle: ", psBlogTitle, 
        "\nauthor: ", psBlogAuthor, 
        "\ndate: ", format(Sys.time(), "%d %B %Y"),
        "\noutput: html_document\n---\n", file = sBlogFilePath)
    cat(" * Created new blog post: ", sBlogFileName, "\n   in directory: ", sBlogFileDir, "\n")
  } else {
    cat(" * Blog post: ", sBlogFileName, "\n   in directory: ", sBlogFileDir, " already exists\n")
  }
  ### # if an image directory was specified, we create it
  if (!is.null(psImgDir)) {
    sImgDir <- file.path(psImgDir, sBlogFileStem)
    if (!dir.exists(sImgDir)) {
      dir.create(sImgDir)
      cat(" * Created Image directory: ", sImgDir, "\n")
    } else {
      cat(" * Image directory: ", sImgDir, " already exists\n")
    }    
  }

}
```

An example call to that function is shown below

```{r CreateGhPagesBlogPostCall}
createGhPagesBlogPost(psBlogTitle  = "Eclipse CDT As Alternative To XCode On The Mac", 
                      psBlogAuthor = "Peter von Rohr", 
                      psBlogName   = "EclipseCdtXcodeAlt", 
                      psRootDir    = "C:/Daten/GitHub/charlotte-ngs.github.io",
                      psImgDir     = "C:/Daten/GitHub/charlotte-ngs.github.io/img")
```
